Magnetized accretion disks around Kerr black holes with scalar hair - I. Constant angular momentum

### 2018-07-30 paper_prd.tex
Some work on the intro.

In recent years, new families of stationary, asymptotically flat black holes (BHs) in general relativity, avoiding the so-called ``no hair" theorems, have been obtained. In particular, Kerr BHs with synchronised hair~\cite{Herdeiro:2014a,Herdeiro:2016} are a counter example to the no hair conjecture, in general relativity minimally coupled to simple (bosonic) matter fields obeying all energy conditions. The physical conditions and stability properties of these {\it hairy} BHs  (HBHs) have been investigated to assess their potential viability as valid alternatives to astrophysical Kerr BHs.  On the one hand, HBHs have been shown to form dynamically as the end-product of the superradiant instability~\cite{East:2017,Herdeiro:2017}. Moreover, the stability of these solutions, questioned by~\cite{Ganchev:2018} in the scalar case, has been recently revisited by~\cite{Degollado:2018} who have shown that the domain of HBH solutions for which the superradiant instability might affect their actual existence has no significance for astrophysical BHs, for the largest timescale possible (the age of the Universe), rendering hairy BHs as {\it effectively} stable.

In addition, the detection of gravitational waves from binary BHs~\cite{Abbott2016, Abbott:2016nmj, Abbott:2017vtc, Abbott:2017oio, Abbott:2017gyy} and the prospect of observing the first image - the black hole {\it shadow} - of a BH by the Event Horizon Telescope~\cite{Fish:2016} opens the opportunity to test the true nature of BHs and the astrophysical relevance of HBHs. 
Regarding the latter, the source of light in a realistic situation is provided by material from an accretion disk crossing the event horizon of the BH, producing the shadow. Thick accretion disks (or tori) are common systems in astrophysics, either surrounding the central BHs of quasars and active galactic nuclei or, at a stellar scale, surrounding the compact objects in X-ray binaries, microquasars, and gamma-ray bursts. Motivated by this observational possibility, in this work we build equilibrium solutions of HBHs surrounded by magnetized accretion disks. Our procedure, which combines earlier approaches  
by~\cite{Komissarov:2006,Qian:2009} was presented in~\cite{Gimeno-Soler:2017} for the Kerr BH case. In~\cite{Gimeno-Soler:2017} we built equilibrium sequences of accretion disks  endowed with a purely toroidal magnetic field, assuming a form of the angular momentum distribution that departs from the constant case considered by~\cite{Komissarov:2006} and from which the location and morphology of the equipotential surfaces can be numerically computed. Our immediate goal in the present work is to assess if and how the accretion disk morphology depends on the type of BH considered, either Kerr BHs of varying spins or Kerr BHs with scalar hair (KBHsSH). Moreover, we focus on disks with a constant distribution of specific angular momentum, presenting in a companion paper the non-constant (power-law) case. In subsequent investigations, we will analyze the shadows of the equilibrium models of KBHsSH presented in this work as well as in {\it dynamical} situations.

>>>>

In recent years, new families of stationary, asymptotically flat black holes (BHs) in general relativity, avoiding the so-called ``no hair" theorems, have been obtained (see e.g.~\cite{Herdeiro:2015_review} and references therein). In particular, Kerr BHs with synchronised hair~\cite{Herdeiro:2014a,Herdeiro:2016} are a counterexample to the no hair conjecture, resulting from minimally coupling Einstein's gravity to simple (bosonic) matter fields obeying all energy conditions. The physical conditions and stability properties of these {\it hairy} BHs  (HBHs) have been investigated to assess their potential viability as  alternatives to astrophysical Kerr BHs.  On the one hand, Kerr BHs with Proca hair have been shown to form dynamically as the end-product of the superradiant instability~\cite{East:2017,Herdeiro:2017} (see also~\cite{Sanchis-Gual:2016,Bosch:2016} for the case of a charged scalar field around a charged BH
in spherical  symmetry). On the other hand, the stability of these solutions, questioned by~\cite{Ganchev:2018} in the scalar case, has been recently revisited by~\cite{Degollado:2018} who have shown that the domain of HBH solutions for which the superradiant instability might affect their actual existence has no significance for astrophysical BHs, rendering HBHs as {\it effectively} stable.

In the observational arena, the LIGO/Virgo detection of gravitational waves from binary BHs~\cite{Abbott2016, Abbott:2016nmj, Abbott:2017vtc, Abbott:2017oio, Abbott:2017gyy} and the exciting prospects of observing the first image - the black hole {\it shadow} - of a BH by the Event Horizon Telescope~\cite{Fish:2016} opens the opportunity to test the true nature of BHs and the astrophysical relevance of HBHs. 
Regarding the latter, the source of light in a realistic situation is provided by material from an accretion disk crossing the event horizon of the BH, producing the shadow. Thick accretion disks (or tori) are common systems in astrophysics, either surrounding the central BHs of quasars and active galactic nuclei or, at a stellar scale, surrounding the compact objects in X-ray binaries, microquasars, and gamma-ray bursts. Motivated by this observational possibility, in this work we build equilibrium solutions of HBHs surrounded by magnetized accretion disks. Our procedure, which combines earlier approaches  
by~\cite{Komissarov:2006,Qian:2009} was presented in~\cite{Gimeno-Soler:2017} for the Kerr BH case. In~\cite{Gimeno-Soler:2017} we built equilibrium sequences of accretion disks  endowed with a purely toroidal magnetic field, assuming a form of the angular momentum distribution that departs from the constant case considered by~\cite{Komissarov:2006} and from which the location and morphology of the equipotential surfaces can be numerically computed. Our immediate goal in the present work is to assess if and how the accretion disk morphology depends on the type of BH considered, either Kerr BHs of varying spins or Kerr BHs with scalar hair (KBHsSH). Moreover, we focus on disks with a constant distribution of specific angular momentum, presenting in a companion paper the non-constant (power-law) case. In subsequent investigations, we will analyze the shadows of the equilibrium models of KBHsSH presented in this work as well as in {\it dynamical} situations.
